---
title: "Introduction to metabasket"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to metabasket}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(metabasket)
```

## What are Basket Trials?

Basket trials are a type of master protocol that evaluates a single therapy across multiple cancer types or subtypes simultaneously. The term "basket" refers to grouping patients with different cancers that share a common molecular alteration or biomarker.

### Key Features

- **Single Treatment**: All patients receive the same investigational therapy
- **Multiple Baskets**: Each basket represents a different cancer type or subtype
- **Information Borrowing**: Statistical methods can borrow information across baskets to improve efficiency
- **Efficiency**: One protocol, one study team, faster enrollment

### Example Trials

1. **Imatinib in Sarcomas** (Chugh et al. 2009): Evaluated imatinib across 10 sarcoma subtypes
2. **Vemurafenib in BRAF V600 Cancers** (Hyman et al. 2015): Tested vemurafenib in 6 non-melanoma cancers
3. **Larotrectinib in TRK Fusion Cancers** (Drilon et al. 2018): Studied TRK inhibitor across 12 cancer types

## The metabasket Package

The `metabasket` package provides a unified framework for:

1. **Designing** basket trials with different statistical methods
2. **Simulating** trial data to evaluate operating characteristics
3. **Analyzing** basket trial data with consistent interfaces
4. **Comparing** different design approaches
5. **Generating** protocol language for regulatory submissions

## System Requirements

Most methods in `metabasket` work out of the box. However, the **Bayesian Hierarchical Model (BHM)** methods require **JAGS** (Just Another Gibbs Sampler) to be installed on your system.

### Installing JAGS

**macOS (using Homebrew):**
```bash
brew install jags
```

After installing JAGS, the `rjags` R package may need symlinks if Homebrew installed JAGS in `/opt/homebrew`:
```bash
sudo mkdir -p /usr/local/lib
sudo ln -sf /opt/homebrew/Cellar/jags/*/lib/libjags.4.dylib /usr/local/lib/libjags.4.dylib
sudo ln -sf /opt/homebrew/Cellar/jags/*/lib/JAGS /usr/local/lib/JAGS
```

Then install the R package:
```r
install.packages("rjags")
```

**Windows:**

1. Download the JAGS installer from: https://mcmc-jags.sourceforge.io/
2. Run the installer (choose appropriate 32-bit or 64-bit version)
3. Install the R package:
```r
install.packages("rjags")
```

**Linux (Ubuntu/Debian):**
```bash
sudo apt-get install jags
```

Then in R:
```r
install.packages("rjags")
```

**Verifying Installation:**
```r
library(rjags)
# Should load without errors
```

If BHM methods are not needed, all other methods (BMA, MEM, Simon, Cunanan) work without JAGS.

## Basic Workflow

### 1. Define a Basket Trial Design

```{r}
# Create a simple 4-basket design
design <- basket_design(
  n_baskets = 4,
  basket_names = c("NSCLC", "SCLC", "Melanoma", "RCC"),
  sample_sizes = 25,
  response_rates = c(0.35, 0.35, 0.35, 0.35),
  null_response_rates = 0.20,
  design_type = "bma"
)

print(design)
```

### 2. Simulate Trial Data

```{r}
# Generate simulated trials
set.seed(123)
simulated_data <- simulate_responses(design, n_sims = 10)

# Look at first simulation
print(simulated_data[[1]])
```

### 3. Analyze Data

```{r, eval=FALSE}
# Note: These functions will be fully implemented with package integrations

# Analyze a single trial
results <- analyze_basket(simulated_data[[1]], design)

# Run full simulation study
sim_results <- simulate_basket_trial(design, n_sims = 1000)

# View operating characteristics
print(sim_results)

# For large simulations, enable parallel processing for speedup
# future::plan(future::multisession, workers = 4)
# sim_results_parallel <- simulate_basket_trial(design, n_sims = 5000, .parallelize = TRUE)
# future::plan(future::sequential)  # Reset when done
```

## Working with Real Data

You can create basket data objects from real trial results:

```{r}
# Imatinib trial data (Chugh et al. 2009)
imatinib_data <- basket_data(
  basket_names = c("Angiosarcoma", "Ewing", "Fibrosarcoma", "Leiomyosarcoma",
                   "Liposarcoma", "MFH", "Osteosarcoma", "MPNST",
                   "Rhabdomyosarcoma", "Synovial"),
  n_patients = c(15, 13, 12, 28, 29, 29, 26, 5, 2, 20),
  n_responses = c(2, 0, 1, 6, 7, 3, 5, 1, 0, 3)
)

print(imatinib_data)
```

## Comparing Design Methods

The package supports multiple statistical approaches:

### Bayesian Model Averaging (BMA)

```{r}
design_bma <- basket_design(
  n_baskets = 4,
  sample_sizes = 25,
  response_rates = c(0.35, 0.35, 0.20, 0.20),
  null_response_rates = 0.20,
  design_type = "bma"
)
```

BMA considers all possible partitions of baskets and averages results across partitions weighted by their posterior probabilities.

### Multi-source Exchangeability Model (MEM)

```{r}
design_mem <- basket_design(
  n_baskets = 4,
  sample_sizes = 25,
  response_rates = c(0.35, 0.35, 0.20, 0.20),
  null_response_rates = 0.20,
  design_type = "mem"
)
```

MEM dynamically learns which baskets are exchangeable based on observed data.

### Bayesian Hierarchical Model (BHM)

```{r}
design_bhm <- basket_design(
  n_baskets = 4,
  sample_sizes = 25,
  response_rates = c(0.35, 0.35, 0.20, 0.20),
  null_response_rates = 0.20,
  design_type = "bhm"
)
```

BHM assumes basket response rates come from a common distribution, enabling shrinkage estimation.

## Generating Protocol Language

The package can generate standardized protocol text:

```{r}
protocol_text <- generate_protocol_language(
  design_bma,
  include_statistical_details = TRUE,
  include_references = TRUE
)

# Display protocol text
cat(protocol_text)
```

```{r, eval=FALSE}
# Export to file
export_protocol_language(design_bma, file = "basket_protocol.md")
```

## Understanding Operating Characteristics

Key metrics for basket trials:

### Basket-Specific Metrics

- **Power**: Probability of correctly identifying a promising basket
- **Type I Error**: Probability of falsely declaring a non-promising basket as promising

### Family-Wise Metrics

- **FWER**: Family-wise error rate (probability of ≥1 false positive)
- **FWP-D**: Disjunctive power (probability of detecting ≥1 promising basket)
- **FWP-C**: Conjunctive power (probability of detecting all promising baskets)

## The Cunanan Efficient Design

The Cunanan et al. (2017) design offers a frequentist alternative with an adaptive two-stage approach:

### Key Features

1. **Interim Heterogeneity Test**: Uses Fisher's exact test to assess homogeneity
2. **Two Design Paths**:
   - **Heterogeneous Path**: Analyze baskets separately with Bonferroni correction
   - **Homogeneous Path**: Pool all baskets for increased power
3. **Sample Size Efficiency**: Can achieve high power with fewer patients when treatment effect is homogeneous

### Example Usage

```{r, eval=FALSE}
# Design for 5 baskets (Cunanan default parameters)
design_cunanan <- basket_design(
  n_baskets = 5,
  sample_sizes = rep(7, 5),  # Stage 1: 7 patients per basket
  null_response_rates = 0.15,
  response_rates = 0.45,
  design_type = "cunanan",
  design_params = list(
    r_s = 1,      # Min responses to continue individual basket
    r_c = 5,      # Min total responses to continue all baskets
    n2 = 15,      # Stage 2 sample size per continuing basket
    gamma = 0.52, # Heterogeneity test threshold
    alpha_s = 0.07, # Separate analysis significance level
    alpha_c = 0.05  # Combined analysis significance level
  )
)

# Stage 1 data (interim analysis)
stage1_data <- basket_data(
  basket_names = paste0("Basket", 1:5),
  n_patients = rep(7, 5),
  n_responses = c(4, 4, 1, 0, 1)
)

# Perform interim analysis
interim_result <- analyze_basket(stage1_data, design_cunanan)
print(interim_result)

# Continues to stage 2 based on interim result...
# After collecting additional data:
# final_data <- basket_data(...)
# final_result <- analyze_basket(final_data, design_cunanan, 
#                                interim_data = interim_result)
```

### When to Use Cunanan Design

**Advantages:**
- Frequentist framework (may be preferred by some reviewers)
- Explicit control of FWER at target level
- Efficient when treatment effect is homogeneous
- Clear decision rules

**Considerations:**
- Less power when effect is in only 1 basket
- Requires two-stage implementation
- Less flexible than Bayesian approaches for complex borrowing

## Simon Two-Stage Design (Reference)

The Simon design serves as a reference comparator, analyzing each basket independently:

### Key Features

1. **No Information Borrowing**: Each basket analyzed separately
2. **Optimal or Minimax**: Choose design that minimizes expected N or maximum N
3. **Futility Stopping**: Two-stage design with interim futility check per basket
4. **FWER Control**: Bonferroni correction across baskets

### Example Usage

```{r, eval=FALSE}
# Parallel Simon designs for 4 baskets
design_simon <- basket_design(
  n_baskets = 4,
  sample_sizes = 25,
  null_response_rates = 0.20,
  response_rates = 0.40,
  design_type = "simon"
)

# Stage 1
stage1 <- basket_data(
  basket_names = paste0("Basket", 1:4),
  n_patients = c(12, 12, 12, 12),
  n_responses = c(6, 2, 5, 1)
)

# Interim analysis with Bonferroni correction
interim <- analyze_basket(stage1, design_simon, 
                         alpha = 0.05/4,  # FWER control
                         beta = 0.20)
print(interim)

# Compare with basket designs
# Basket designs typically require fewer patients due to information borrowing
```

### When to Use Simon Design

**Use as reference when:**
- Comparing efficiency of basket designs
- Conservative approach needed (no borrowing assumptions)
- Strong heterogeneity expected (borrowing inappropriate)

**Limitations:**
- Requires more patients than basket designs when treatment is homogeneous
- No adaptation based on similarity across baskets
- Conservative (may miss opportunities for information sharing)

## Design Considerations

### When to Borrow Information

Information borrowing is beneficial when:

- Treatment response rates are expected to be similar across baskets
- Some baskets have small sample sizes
- The mechanism of action suggests similar efficacy

Be cautious when:

- Substantial heterogeneity is expected
- Risk of inflated Type I error is unacceptable
- Strong prior evidence suggests differential effects

### Choosing Sample Sizes

Factors to consider:

1. Expected response rates under H0 and H1
2. Desired power and Type I error control
3. Degree of information borrowing
4. Prevalence of each cancer type
5. Enrollment feasibility

### Setting Null Response Rates

The null rate (p0) typically represents:

- Historical response rate to standard of care
- Minimum clinically interesting response rate
- May vary by basket if standard of care differs

## Next Steps

For more details, see these vignettes:

- `vignette("simulation-studies")`: Conducting simulation studies
- `vignette("method-comparison")`: Comparing statistical methods
- `vignette("real-data-analysis")`: Analyzing published basket trials

## References

1. **Zhou T, Ji Y.** Bayesian Methods for Information Borrowing in Basket Trials: An Overview. *Cancers*. 2024;16(2):251.

2. **Hobbs BP, et al.** Basket trials: Review of current practice and innovations for future trials. *J Clin Oncol*. 2022;40(28):3520-3528.

3. **Psioda MA, et al.** Bayesian adaptive basket trial design using model averaging. *Biostatistics*. 2021;22(1):19-34.

4. **Hobbs BP, Landin R.** Bayesian basket trial design with exchangeability monitoring. *Stat Med*. 2018;37(25):3557-3572.

5. **Berry SM, et al.** Bayesian hierarchical modeling of patient subpopulations: Efficient designs of phase II oncology clinical trials. *Clin Trials*. 2013;10(5):720-734.
